---
title: "Class05_hwQ6_Darin"
author: "Emily Darin"
format: pdf
---
Loading R package bio3d

```{r}
library(bio3d)
```

```{r}
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```


>What type of object is returned from the read.pdb() function?

A summary of infomation PDB file with 8 elements. 

>Q2. What does the trim.pdb() function do?

Makes a smaller pdb object with a subset of atoms from the larger pdb object

>3 What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?

"sse=NULL" would turn off the black a grey rectangles and the represent the protein chain. 
 
```{r}
plotb3(s1.b, sse=NULL, typ="l", ylab="Bfactor")
```
 
>Q4. What would be a better plot to compare across the different proteins?

A cluster dendrogram would probably be better because it would show you how related the structure of the idfferent proteins are. 

>Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?

S1 and S3 are more similar to eachother than s2 for the b factor chains

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

>Q6. How would you generalize the original code above to work with any set of input
protein structures?

First I would create my function that would allow this to be made using any protein. 

```{r}

 pdb <- function(X){
 trim.pdb(X, chain="A", elety="CA") #trims pdb 
 X$atom$b #Calls atoms 
 plotb3(X, sse=s1.chainA, typ="l", ylab="Bfactor") #plots 
  }






```


